---
title: "Multivariate HW4"
author: '106070020'
date: "2021年4月26日"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1

#### (a)
```{r}
#Q1(a)
dat <- as.matrix(read.table("C:/Users/eva/Desktop/作業 上課資料(清大)/大四下/多變量/hw4/T6_1.dat"), header=T)
BOD_1<-log(as.numeric(dat[2:12,1]))
SS_1<-log(as.numeric(dat[2:12,2]))
BOD_2<-log(as.numeric(dat[2:12,3]))
SS_2<-log(as.numeric(dat[2:12,4]))
data<-cbind(BOD_1,SS_1,BOD_2,SS_2)
data<-as.data.frame(data)
head(data)
d1<-data$BOD_1-data$BOD_2
d2<-data$SS_1-data$SS_2
diff<-cbind(d1,d2)
d1bar<-mean(d1)
d2bar<-mean(d2)
d<-rbind(d1bar,d2bar)
d #mean
delta<-rep(0,2)
s1<-c(var(d1),cov(d1,d2))
s2<-c(cov(d1,d2),var(d2))
sd<-rbind(s1,s2)
sd #variance matrix
sdt<-solve(sd)
sdt #(sd) inverse
t2<-(t(d)-delta)%*%sdt%*%d
nt2<-nrow(data)*t2
nt2 # T square
#taking alpha=.05
n<-nrow(data)
p<-2
F_stat<-(2*(n-1)/(n-p))*qf(.95, df1=2, df2=9)
F_stat #F score
nt2>F_stat
del1<-c(d1bar-sqrt((n-1)*p/(n-p)*qf(.95, df1=2, df2=9)*var(d1)/n),d1bar+sqrt((n-1)*p/(n-p)*qf(.95, df1=2, df2=9)*var(d1)/n))
del2<-c(d2bar-sqrt((n-1)*p/(n-p)*qf(.95, df1=2, df2=9)*var(d2)/n),d2bar+sqrt((n-1)*p/(n-p)*qf(.95, df1=2, df2=9)*var(d2)/n))
del1 #CI of delta1
del2 #CI of delta2
```

***

+ Since T^2=10.21541 > F score=9.458877, we can reject H0 and conclude that there is a nonzero mean difference between the mesurements of two labortories.
+ The 95% simutaneous intervals for the mean difference delta are:
    + -1.1844404 < delta1 < 0.0680501
    + -0.1020988 < delta2 < 0.6931554

#### (b)
```{r}
t_10<-qt(1-0.05/4,10)
t_10 
bdel1<-c(d1bar-t_10*sqrt(var(d1)/n),d1bar+t_10*sqrt(var(d1)/n))
bdel2<-c(d2bar-t_10*sqrt(var(d2)/n),d2bar+t_10*sqrt(var(d2)/n))
bdel1
bdel2
```

***
*  The 95% Bonferroni simutaneous intervals for the components of the mean vector delta of transformed variables are:
    + -1.09448795 < delta1 <-0.02190232
    + -0.04498455 < delta2 < 0.63604112

#### (c)
```{r message=F}
#(c)
library(MVN)
result<-mvn(diff, mvnTest = 'mardia')
result$multivariateNormality
result<-mvn(diff, mvnTest = 'hz')
result$multivariateNormality
```

***
+ Since both tests indicate multivariate normality, then data diff follows a multivariate normalitydistribution at the 0.05 confidence level.

## Problem 2

#### (a)
```{r}
#Q2(a)
treatment2<-rbind(c(3,3),c(1,6),c(2,3))
treatment3<-rbind(c(2,3),c(5,1),c(3,1), c(2,3))

mtr2<-c(mean(treatment2[,1]),mean(treatment2[,2]))
mtr2
s11<-var(treatment2[,1])
s12<-var(treatment2[,1],treatment2[,2])
s22<-var(treatment2[,2])
str2<-rbind(c(s11,s12),c(s12,s22))
str2
mtr3<-c(mean(treatment3[,1]),mean(treatment3[,2]))
mtr3
s311<-var(treatment3[,1])
s312<-var(treatment3[,1],treatment3[,2])
s322<-var(treatment3[,2])
str3<-rbind(c(s311,s312),c(s312,s322))
str3
s_pool<-(3-1)/(3+4-2)*str2+(4-1)/(3+4-2)*str3
s_pool
```

#### (b)
```{r}
#(b)
T_sqr<-t(mtr2-mtr3)%*%solve((1/3+1/4)*s_pool)%*%(mtr2-mtr3)
T_sqr
c_sqr<-(3+4-2)/(3+4-2-1)*2*qf(.99, 2, (3+4-2-1))
c_sqr
T_sqr < c_sqr
```

***
+ H0: The difference of mu1 and mu2 is 0. H1:The difference of mu1 and mu2 is not 0.
+ T_sqr=3.870968 < c_sqr=45, so we cannot reject H0.

#### (c)
```{r}
#i=1
ci1<-c((mtr2[1]-mtr3[1])-sqrt(c_sqr*(1/3+1/4)*s_pool[1,1]),(mtr2[1]-mtr3[1])+sqrt(c_sqr*(1/3+1/4)*s_pool[1,1]))
#i=2
ci2<-c((mtr2[2]-mtr3[2])-sqrt(c_sqr*(1/3+1/4)*s_pool[2,2]),(mtr2[2]-mtr3[2])+sqrt(c_sqr*(1/3+1/4)*s_pool[2,2]))
ci1
ci2
```

***
> 99% simultaneous CI for the difference :

+ -7.480741 < mu11-mu21 < 5.480741
+ -5.245688 < mu12-mu22 < 9.245688

## Problem 3

#### (a)
```{r}
data3<-read.csv("C:/Users/eva/Desktop/作業 上課資料(清大)/大四下/多變量/hw4/problem3.csv")
own<-data3$CONTROL
con1<-data3$LATITUDE
con2<-data3$LONGITUDE
uu<-lm(cbind(con1, con2) ~ own)
anova(uu)
summary.aov(uu)
```

***
> 
+ H0: The location of school does not differs among different ownership.
+ H1: The location of school does differs among different ownership.
+ F statistic = 0.1733 > 0.05
+ Conclusion : Cannot reject H0

#### (b)
```{r message=F}
# 1. Public
# 2. Private nonprofit
# 3. Private for-profit
# 4. Foreign

library(dplyr)
ownloc<-cbind(own,con1,con2)
ownloc=as.data.frame(ownloc)
public<-filter(.data=ownloc, ownloc$own == "1")
private<-filter(.data=ownloc, ownloc$own  %in% c("2","3"))

p1<-data.frame(type=rep(1,nrow(public)),latitude=public$con1, longtitude=public$con2)
p2<-data.frame(type=rep(2,nrow(private)),latitude=private$con1, longtitude=private$con2)
pmd<-rbind(p1,p2)
head(pmd)
tvart<-t.test(cbind(pmd$latitude,pmd$longtitude)~pmd$type , var.equal = T)
tvart
tvarf<-t.test(cbind(pmd$latitude,pmd$longtitude)~pmd$type , var.equal = F)
tvarf
```

***
>
+ H0 : True difference in means is equal to 0
+ H1 : True difference in means is not equal to 0

+ Testing statistics: 
    + With equal variance: t = -0.48158, df = 238, p-value = 0.6305
    + Without equal variance: t = -0.47715, df = 154.28, p-value = 0.6339
+ We cannot reject H0 with or without the equal variance. 

#### (c)
```{r}
library(ellipse)
#(c)
public<-filter(.data=ownloc, ownloc$own == "1")
private<-filter(.data=ownloc, ownloc$own  %in% c("2","3"))

mpub<-c(mean(public$con1),mean(public$con2))
pubs11<-var(public$con1)
pubs12<-var(public$con1,public$con2)
pubs22<-var(public$con2)
pubstr2<-rbind(c(pubs11,pubs12),c(pubs12,pubs22))

mpri<-c(mean(private$con1),mean(private$con2))
pris311<-var(private$con1)
pris312<-var(private$con1,private$con2)
pris322<-var(private$con2)
pristr3<-rbind(c(pris311,pris312),c(pris312,pris322))

pps_pool<-(40-1)/(40+80-2)*pubstr2+(80-1)/(40+80-2)*pristr3
d.bar<-mpub-mpri
conf.ellipse<-ellipse(pps_pool,centre=d.bar, level=0.95)
{plot(conf.ellipse, type="l", main='The difference of location centers between public schools
      and non-public schools', lwd=2, col='blue')
points(x=d.bar[1],y=d.bar[2],pch=16, col="red")
text(x=-2, y=1, 'The difference of location centers', col='red')}

```

***
> 
+ The result consistent with the output of (b), as the difference of location centers is inside the 95% confidence ellipse for the difference of location centers between public schools and non-public schools. + Conclusion: We cannot reject H0.(The schools of the two groups aren't distributed at different 
locations.)

#### (d)
```{r message=F}
library(knitr)
X1<-colMeans(public[,2:3])
X2<-colMeans(private[,2:3])
S1<-var(public[,2:3])
S2<-var(private[,2:3])
n1<-nrow(public)
n2<-nrow(private)
Sp<-((n1-1)*S1+(n2-1)*S2)/(n1+n2-2)
p<-dim(public[,2:3])[2]
delta0 <- rep(0,p)
T2 <- (n1*n2/(n1+n2))*t(X1-X2-delta0)%*%solve(Sp)%*%(X1-X2-delta0)
crit <- (n1 + n2 - 2)*p/(n1 + n2 - p - 1)*qf(1-.05, p, n1 + n2 - p - 1)
A<-diag(p)
T2 > crit
lower <- t(A)%*%(X1 - X2) - sqrt(crit) * sqrt(diag((t(A)%*%Sp%*%A)*(1/n1 + 1/n2)))
upper <- t(A)%*%(X1 - X2) + sqrt(crit) * sqrt(diag((t(A)%*%Sp%*%A)*(1/n1 + 1/n2)))
out <- as.data.frame(cbind(lower, upper), 
                     row.names = c("Lontitude", "Latitude"))
names(out) <- c("Lower", "Upper")
kable(out)

```

## Problem 4

#### (a)(i) Fit appropriate linear model
```{r message=F}
library(survMisc)
library(dplyr)

#(i)
ami<-read.table("C:/Users/eva/Desktop/作業 上課資料(清大)/大四下/多變量/hw4/T7_6.dat", header=T)
head(ami)
fit.y1<-lm(y1~factor(z1)+z2+z3+z4+z5,data=ami)
summary(fit.y1)
c.y1 <- round(fit.y1$coefficients, digits = 4)
c.y1
anova(fit.y1)

```

***
> MODEL: Y1=−2879.4782 + 675.6508z1 + 0.2849z2 + 10.2721z3 + 7.2512z4 + 7.5982z5

#### (a)(ii) Residual Analysis
```{r message=F, warning=FALSE, comment=FALSE}
#(ii)
library(ggfortify)
autoplot(fit.y1, which = 1:6, ncol = 3, label.size = 3)
```


#### (a)(iii)Prediction Interval
```{r}
#(iii)
pred.y1 <- data.frame(z1 = 1, z2 = 1200, z3 = 140, z4 = 70, z5 = 85)
PI.95.y1 <- predict(fit.y1, newdata = pred.y1, interval = 'prediction')
dimnames(PI.95.y1)[[2]] <- list("Estimate", "Lower Limit", "Upper Limit")
PI.95.y1
```

#### (b)(i) Fit appropriate linear model
```{r}
fit.y2<-lm(y2~factor(z1)+z2+z3+z4+z5,data=ami)
summary(fit.y2)
c.y2 <- round(fit.y2$coefficients, digits = 4)
c.y2
anova(fit.y2)
```

***
> MODEL: Y2= -2728.7085 + 763.0298z1 + 0.3064z2 + 8.8962z3 + 7.2056z4 + 4.9871z5

#### (b)(ii) Residual Analysis
```{r message=F, warning=FALSE, comment=FALSE}
autoplot(fit.y2, which = 1:6, ncol = 3, label.size = 3)
```

#### (b)(iii) Prediction Interval
```{r}
pred.y2 <- data.frame(z1 = 1, z2 = 1200, z3 = 140, z4 = 70, z5 = 85)
PI.95.y2 <- predict(fit.y2, newdata = pred.y2, interval = 'prediction')
dimnames(PI.95.y2)[[2]] <- list("Estimate", "Lower Limit", "Upper Limit")
PI.95.y2
```
 
####(c)(i) Fit appropriate linear model
```{r message=F}
library(car)
fit.y1y2 <- lm(as.matrix(ami[,1:2]) ~ factor(z1) + z2 + z3 + z4 + z5 , data = ami)
y1y2.sum <- summary(fit.y1y2)
y1y2.sum
c.y1y2 <- round(fit.y1y2$coefficients, digits = 4)
c.y1y2

man.y1y2 <- Manova(fit.y1y2)
summary(man.y1y2, "Wilks")
```

#### (c)(ii) Residual Analysis
```{r message=F, warning=FALSE, comment=FALSE}
plot(fit.y1y2$residuals)
```

#### (c)(iii) Prediction Interval
```{r message=F, warning=FALSE, comment=FALSE}
pred.y1y2 <- data.frame(z1 = 1, z2 = 1200, z3 = 140, z4 = 70, z5 = 85)
point <- as.data.frame(predict(fit.y1y2, newdata = pred.y1y2, interval = 'prediction'))
center <- c(point[1,1], point[1,2])

Z <- model.matrix(fit.y1y2)
resp <- fit.y1y2$model[[1]]#list(if 1[], type will be matrix)
n <- nrow(resp); m <- ncol(resp); r <- ncol(Z) - 1
S <- crossprod(fit.y1y2$residuals)/(n - r - 1)#inner product

t2 <- terms(fit.y1y2)
term <- delete.response(t2)
mframe <- model.frame(term, pred.y1y2, na.action = na.pass, xlev = fit.y1y2$xlevels)
z0 <- model.matrix(term, mframe, contrasts.arg = fit.y1y2$contrasts)
radius <- sqrt((m*(n - r - 1)/(n - r - m))*qf(0.95, m, n - r - m)*z0%*%solve(t(Z)%*%Z) %*% t(z0))

lipsy <-as.data.frame(ellipse(center = c(center), shape = S, radius = c(radius), draw = FALSE))


ggplot(lipsy, aes(x, y)) + 
  geom_path() + 
  geom_point(aes(x = y1, y = y2), data = point, alpha = 0.5) + 
  labs(x = "Total TCAD plasma level (plasma)", 
       y = "Amount of amitriptyline present in TCAD plasma level (ami)", 
       title = "95% Prediction Ellipse for Given Predictor Setting")

ggplot(lipsy, aes(x, y)) + 
  geom_hline(aes(y = PI.95.y1[[2]]), 
             yintercept = PI.95.y1[[2]], 
             color = "blue", 
             linetype = "dashed") + 
  geom_hline(aes(y = PI.95.y1[[3]]), 
             yintercept = PI.95.y1[[3]], 
             color = "blue", 
             linetype = "dashed") + 
  geom_vline(aes(x = PI.95.y2[[2]]), 
             xintercept = PI.95.y2[[2]], 
             color = "red", 
             linetype = "dashed") + 
  geom_vline(aes(x = PI.95.y2[[3]]), 
             xintercept = PI.95.y2[[3]], 
             color = "red", 
             linetype = "dashed") + 
  geom_path() + 
  geom_point(aes(x = y1, y = y2), 
             data = point, 
             alpha = 0.5) + 
  labs(x = "Total TCAD plasma level (plasma)", 
       y = "Amount of amitriptyline present in TCAD plasma level (ami)", 
       title = "95% Prediction Ellipse for Given Predictor Setting",
       subtitle = "Red/Blue Dashed Lines Represent Prediction Interval Limits from Parts (a) & (b)") 
```













